---
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---

```{r intro, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      comment = "", 
                      warning=FALSE, 
                      results="hide") 

knitr::opts_knit$set(root.dir = "C:/Users/YoonJoung Choi/Dropbox/0 iSquared/iSquared_PMA/Effective Access/")

date<-as.Date(Sys.time(	), format='%d%b%Y')
time<-Sys.time()

library(plyr)
library(dplyr)
library(tidyverse)
library(plotly)
library(lubridate)

# Libraries for radar chart
library(viridis)
library(patchwork)
library(hrbrthemes)
library(fmsb)
library(colormap)

```

```{r dta, results='hide'}
dtaIR<-data.frame(read_csv("summary_Access_Indicators_IR.csv"))%>%
    mutate(yearmonth = ISOdate(year, month, 15),
           #yearmonth = as.POSIXct(as.numeric(yearmonth), origin="1970-01-01"), 
           #yearmonth = substr(as.character(yearmonth),0,7), 
           latestIRLINK=latestIR,
           latestIRLINK=ifelse(xsurvey=="BFR5", 1, latestIRLINK),
           latestIRLINK=ifelse(xsurvey=="BFR6", 0, latestIRLINK),
           latestIRLINK=ifelse(xsurvey=="BFR7", 0, latestIRLINK),
           latestIRLINK=ifelse(xsurvey=="KER7", 1, latestIRLINK),
           latestIRLINK=ifelse(xsurvey=="KER8", 0, latestIRLINK), 
           
           linkissue=(country=="Nigeria, Kano" | country=="Nigeria, Lagos" | country=="India, Rajasthan" | 
                      xsurvey=="NGKanoR3" | xsurvey=="BFR6"),
           SDPall_essential5_rnoso=ifelse(linkissue==1, NA, SDPall_essential5_rnoso), 
           SDPall_essential5_noso=ifelse(linkissue==1, NA, SDPall_essential5_noso), 
           SDPall_essential5_ready=ifelse(linkissue==1, NA, SDPall_essential5_ready), 
           SDPall_essential5ec_rnoso=ifelse(linkissue==1, NA, SDPall_essential5_rnoso), 
           SDPall_essential5ec_noso=ifelse(linkissue==1, NA, SDPall_essential5_noso), 
           SDPall_essential5ec_ready=ifelse(linkissue==1, NA, SDPall_essential5_ready) 
           
    )

dtaCR<-data.frame(read_csv("summary_Access_Indicators_CR.csv"))%>%
    mutate(yearmonth = ISOdate(year, month, 15)
           #yearmonth = substr(as.character(yearmonth),0,7)
    )

    #table(dtaIR$xsurvey)
    #table(dtaCR$xsurvey)
    
    #table(dtaIR$group)
    #table(dtaCR$group)
    
    #table(dtaIR$grouplabel)
    #table(dtaCR$grouplabel)    

    #table(dtaIR$yearmonth)
    #table(dtaCR$yearmonth)

dta<-rbind.fill(dtaIR, dtaCR)
```

(last updated: `r time`)

__Potential data viz exploration for women's access to family planning programs__ 

Objectives:    
1. communicate multi-dimensional access. The bigger/more round web, the better/more balanced access. (Note: Radar chart is useful specifically for XYZ)           
2. visualize how this multi-dimensional access varies by sub-group.     

Caveats:    
1. In each dimension, a potential candidate metric is used. Part of this exercise is to identify/revise the candidates. So, they are not final. A few domains still have relatively poor data. Nevertheless, the metrics used here are:   

* FP.agency: percent of women who strongly agreed or agreed with each of the following five statements:    
* FP.efficacy: percent of women who strongly agreed or agreed with each of the following two statements:    
* Information: percent of women who ever heard of specific 5 methods: IUD, implant, injectables, pills, and male condoms    
* Affordability.insurance: percent of women who have health insurance (pretty bad...)   
* Method.choice: percent of women who live in a community with at least on SDPs that have the five methods + have equipment/personnel to insert/remove IUD and implant + had no history of stock out in the past three months    
* Service quality 1: percent of _FP clients_ who reported the communication was clear, they were allowed to ask questions, and they understood (index based on three questions).    
* Service quality 2: percent of _FP clients_ who reported her provider discussed advantage/disadvantage of the method [she adopted].   

__Questions__    
1. Is it potentially useful?   
2. Too simple to have one metric in each domain? We can have 2 metrics per domain. Not more than that.    
3. 

####__1. Among all women__

```{r dtaRadar, results='hide'}
dtafig<-dta%>%
    filter(xsurvey=="BFR7" | xsurvey=="KER8")%>%
    filter(group=="All" | group=="All clients")%>%
    mutate(
        FP.empowerment =xwge_fp_all, 
        FP.agency=xwge_fp_existence_all,
        FP.efficacy =xwge_fp_exercise_all,
        Information    =xheard_select5, 
        Affordability.insurance   = xinsurance,
        Method.choice=SDPall_essential5_ready, 
        Quality1 = ycommunication_all,
        Quality2 = ydiscuss_procon
    )%>%
    select(xsurvey, 
           FP.empowerment, 
           FP.agency, 
           FP.efficacy, 
           Information, 
           Affordability.insurance, 
           Method.choice, 
           Quality1,
           Quality2)

    temp<-dtafig[,-1]

dtafig<-dtafig%>%
    group_by(xsurvey)%>%
    summarise_at(c(colnames(temp)), max, na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        Method.choice=ifelse(xsurvey=="KER8", 73, Method.choice),
        Method.choice=ifelse(xsurvey=="BFR7", 91, Method.choice) 
    )
    #pop-level access values borrowed from past survey in the country. "link to any SDPs": https://rpubs.com/YJ_Choi/AccessToMethods

grouplist<-c(dtafig$xsurvey)
grouplist

maxvalue<-c(" ", rep(100,8))
minvalue<-c(" ", rep(0,8))

dtaradar <- rbind(maxvalue, minvalue, dtafig)%>%
    select(-xsurvey)%>%
    mutate_if(is.character,as.numeric)

dtaradar<-as.data.frame(dtaradar)

#View(dtafig)
#View(dtaradar)

```

```{r plotRadar, results='asis', fig.align="left", out.width="1000px", out.height="800px"}
# Set screen
par(mar=c(rep(1,4)) )
par(mfrow=c(1,2))

# Loop for each plot
grouplist<-c(dtafig$xsurvey)

for(i in 1:length(grouplist)){

radarchart(dtaradar[c(1,2,i+2), -1], axistype=1,
           
          #custom polygon
          pcol=rgb(0.09,0.36,0.54, 0.8) , 
          pfcol=rgb(0.09,0.36,0.54, 0.3) , 
          plwd=2 , 
        
          #custom the grid
          cglcol="grey", cglty=1, 
          axislabcol="grey", 
          cglwd=0.8,
          caxislabels=seq(0,100,25), 
        
          #custom labels
          vlcex=0.7,
    
          #title
          title=grouplist[i]
          )
}
          
```



####__2. By women's background characteristics: education as an example__

```{r dtaRadar_ByEdu}
dtafig<-dta%>%
    filter(xsurvey=="BFR7" | xsurvey=="KER8")%>%
    filter(group=="Education" | group=="Education (CR)" )%>%
    mutate(grouplist=paste(xsurvey, grouplabel))%>%
    mutate(
        FP.empowerment =xwge_fp_all, 
        FP.agency=xwge_fp_existence_all,
        FP.efficacy =xwge_fp_exercise_all,
        Information    =xheard_select5, 
        Affordability.insurance   = xinsurance,
        Method.choice   =SDPall_essential5_ready, 
        Quality1 = ycommunication_all,
        Quality2 = ydiscuss_procon
    )%>%
    select(grouplist,
           FP.empowerment, 
           FP.agency, 
           FP.efficacy, 
           Information, 
           Affordability.insurance, 
           Method.choice, 
           Quality1,
           Quality2)

    temp<-dtafig[,-1]

dtafig<-dtafig%>%
    group_by(grouplist)%>%
    summarise_at(c(colnames(temp)), max, na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        Method.choice=ifelse(substr(grouplist, 1, 4)=="KER8", 73, Method.choice),
        Method.choice=ifelse(substr(grouplist, 1, 4)=="BFR7", 91, Method.choice)
    )
    #pop-level access values borrowed from past survey in the country. "link to any SDPs": https://rpubs.com/YJ_Choi/AccessToMethods

maxvalue<-c(" ", rep(100,8))
minvalue<-c(" ", rep(0,8))

dtaradar <- rbind(maxvalue, minvalue, dtafig)%>%
    select(-grouplist)%>%
    mutate_if(is.character,as.numeric)

#View(dtafig)
#View(dtaradar)
```

```{r plotRadar_ByEdu, results='asis', fig.align="left", out.width="1000px", out.height="800px"}
# Set screen
par(mar=c(rep(1,4)) )
par(mfrow=c(2,2))

# Loop for each plot
grouplist<-c(dtafig$grouplist)

for(i in 1:length(grouplist)){

radarchart(dtaradar[c(1,2,i+2), -1], axistype=1,
           
          #custom polygon
          pcol=rgb(0.09,0.36,0.54, 0.8) , 
          pfcol=rgb(0.09,0.36,0.54, 0.3) , 
          plwd=2 , 
        
          #custom the grid
          cglcol="grey", cglty=1, 
          axislabcol="grey", 
          cglwd=0.8,
          caxislabels=seq(0,100,25), 
        
          #custom labels
          vlcex=0.7,
    
          #title
          title=grouplist[i]
          )
}
          
```

####__3. By women's FP demand, contraceptive use, and sexual activity status__

Note: 
* Service quality is not included.    

#####__Among women with demand for FP__ 
```{r dtaRadar_ByDenominator1}
dtafig<-dta%>%
    filter(xsurvey=="BFR7" | xsurvey=="KER8")%>%
    filter(group=="By xdenominator" )%>%
    filter(substr(grouplabel,1,6)=="demand" )%>%
    mutate(grouplist=paste(substr(xsurvey,1,2), grouplabel))%>%
    mutate(
        FP.empowerment =xwge_fp_all, 
        FP.agency=xwge_fp_existence_all,
        FP.efficacy =xwge_fp_exercise_all,
        Information    =xheard_select5, 
        Affordability.insurance   = xinsurance,
        Method.choice   =SDPall_essential5_ready
    )%>%
    select(grouplist,
           FP.empowerment, 
           FP.agency, 
           FP.efficacy, 
           Information, 
           Affordability.insurance, 
           Method.choice)

    temp<-dtafig[,-1]

dtafig<-dtafig%>%
    group_by(grouplist)%>%
    summarise_at(c(colnames(temp)), max, na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        Method.choice=ifelse(substr(grouplist, 1, 2)=="KE", 73, Method.choice),
        Method.choice=ifelse(substr(grouplist, 1, 2)=="BF", 91, Method.choice)
    )
    #pop-level access values borrowed from past survey in the country. "link to any SDPs": https://rpubs.com/YJ_Choi/AccessToMethods

maxvalue<-c(" ", rep(100,6))
minvalue<-c(" ", rep(0,6))

dtaradar <- rbind(maxvalue, minvalue, dtafig)%>%
    select(-grouplist)%>%
    mutate_if(is.character,as.numeric)

#View(dtafig)
#View(dtaradar)
```

```{r plotRadar_ByDenominator1, results='asis', fig.align="left", out.width="1000px", out.height="800px"}
# Set screen
par(mar=c(rep(1,4)) )
par(mfrow=c(2,2))

# Loop for each plot
grouplist<-c(dtafig$grouplist)
grouplist<-c("BF demand, use", "BF demand, unmet", 
             "KE demand, use", "KE demand, unmet")
grouplist<-c(dtafig$grouplist)

for(i in 1:length(grouplist)){

radarchart(dtaradar[c(1,2,i+2), -1], axistype=1,
           
          #custom polygon
          pcol=rgb(0.09,0.36,0.54, 0.8) , 
          pfcol=rgb(0.09,0.36,0.54, 0.3) , 
          plwd=2 , 
        
          #custom the grid
          cglcol="grey", cglty=1, 
          axislabcol="grey", 
          cglwd=0.8,
          caxislabels=seq(0,100,25), 
        
          #custom labels
          vlcex=0.7,
    
          #title
          title=grouplist[i]
          )
}
          
```

#####__Among women without demand for FP__
```{r dtaRadar_ByDenominator2}
dtafig<-dta%>%
    filter(xsurvey=="BFR7" | xsurvey=="KER8")%>%
    filter(group=="By xdenominator" )%>%
    filter(substr(grouplabel,1,6)!="demand" )%>%
    mutate(grouplist=paste(substr(xsurvey,1,2), grouplabel))%>%
    mutate(
        FP.empowerment =xwge_fp_all, 
        FP.agency=xwge_fp_existence_all,
        FP.efficacy =xwge_fp_exercise_all,
        Information    =xheard_select5, 
        Affordability.insurance   = xinsurance,
        Method.choice   =SDPall_essential5_ready
    )%>%
    select(grouplist,
           FP.empowerment, 
           FP.agency, 
           FP.efficacy, 
           Information, 
           Affordability.insurance, 
           Method.choice)

    temp<-dtafig[,-1]

dtafig<-dtafig%>%
    group_by(grouplist)%>%
    summarise_at(c(colnames(temp)), max, na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        Method.choice=ifelse(substr(grouplist, 1, 2)=="KE", 73, Method.choice),
        Method.choice=ifelse(substr(grouplist, 1, 2)=="BF", 91, Method.choice)
    )
    #pop-level access values borrowed from past survey in the country. "link to any SDPs": https://rpubs.com/YJ_Choi/AccessToMethods

maxvalue<-c(" ", rep(100,6))
minvalue<-c(" ", rep(0,6))

dtaradar <- rbind(maxvalue, minvalue, dtafig)%>%
    select(-grouplist)%>%
    mutate_if(is.character,as.numeric)

#View(dtafig)
#View(dtaradar)
```

```{r plotRadar_ByDenominator2, results='asis', fig.align="left", out.width="1000px", out.height="800px"}
# Set screen
par(mar=c(rep(1,4)) )
par(mfrow=c(2,3))

# Loop for each plot
grouplist<-c(dtafig$grouplist)
grouplist<-c(
             "BF no demand, fecund", "BF no demand, infecund", 
             "BF not sexually active",
             
             "KE no demand, fecund", "KE no demand, infecund",
             "KE not sexually active")
grouplist<-c(dtafig$grouplist)

for(i in 1:length(grouplist)){

radarchart(dtaradar[c(1,2,i+2), -1], axistype=1,
           
          #custom polygon
          pcol=rgb(0.09,0.36,0.54, 0.8) , 
          pfcol=rgb(0.09,0.36,0.54, 0.3) , 
          plwd=2 , 
        
          #custom the grid
          cglcol="grey", cglty=1, 
          axislabcol="grey", 
          cglwd=0.8,
          caxislabels=seq(0,100,25), 
        
          #custom labels
          vlcex=0.7,
    
          #title
          title=grouplist[i]
          )
}
          
```

---

```{r dtaRadar_ByDenominator}
dtafig<-dta%>%
    filter(xsurvey=="BFR7" | xsurvey=="KER8")%>%
    filter(group=="By xdenominator" )%>%
    mutate(grouplist=paste(substr(xsurvey,1,2), grouplabel))%>%
    mutate(
        FP.empowerment =xwge_fp_all, 
        FP.agency=xwge_fp_existence_all,
        FP.efficacy =xwge_fp_exercise_all,
        Information    =xheard_select5, 
        Affordability.insurance   = xinsurance,
        Method.choice   =SDPall_essential5_ready
    )%>%
    select(grouplist,
           FP.empowerment, 
           FP.agency, 
           FP.efficacy, 
           Information, 
           Affordability.insurance, 
           Method.choice)

    temp<-dtafig[,-1]

dtafig<-dtafig%>%
    group_by(grouplist)%>%
    summarise_at(c(colnames(temp)), max, na.rm = TRUE)%>%
    ungroup()%>%
    mutate(
        Method.choice=ifelse(substr(grouplist, 1, 2)=="KE", 73, Method.choice),
        Method.choice=ifelse(substr(grouplist, 1, 2)=="BF", 91, Method.choice)
    )
    #pop-level access values borrowed from past survey in the country. "link to any SDPs": https://rpubs.com/YJ_Choi/AccessToMethods

maxvalue<-c(" ", rep(100,6))
minvalue<-c(" ", rep(0,6))

dtaradar <- rbind(maxvalue, minvalue, dtafig)%>%
    select(-grouplist)%>%
    mutate_if(is.character,as.numeric)

#View(dtafig)
#View(dtaradar)
```

```{r plotRadar_ByDenominator, eval=FALSE}
# Set screen
par(mar=c(rep(1,4)) )
par(mfrow=c(2,5))

# Loop for each plot
grouplist<-c(dtafig$grouplist)
grouplist<-c("BF demand, use", "BF demand, unmet", 
             "BF no demand, fecund", "BF no demand, infecund", 
             "BF not sexually active",
             "KE demand, use", "KE demand, unmet",
             "KE no demand, fecund", "KE no demand, infecund",
             "KE not sexually active")
grouplist<-c(dtafig$grouplist)

for(i in 1:length(grouplist)){

radarchart(dtaradar[c(1,2,i+2), -1], axistype=1,
           
          #custom polygon
          pcol=rgb(0.09,0.36,0.54, 0.8) , 
          pfcol=rgb(0.09,0.36,0.54, 0.3) , 
          plwd=2 , 
        
          #custom the grid
          cglcol="grey", cglty=1, 
          axislabcol="grey", 
          cglwd=0.8,
          caxislabels=seq(0,100,25), 
        
          #custom labels
          vlcex=0.7,
    
          #title
          title=grouplist[i]
          )
}
          
```


---


```{r plotlyRadar, fig.align="left", out.width="1000px", out.height="800px"}

figBF<-plot_ly(
    type = 'scatterpolar',
    r = c(38,43,47,2,91,33,30,38),
    theta = c("FP.agency", "FP.efficacy", 
              "Information", "Affordability.insurance", "Method.choice",
              "Quality1", "Quality2",
              "FP.agency"),
    name="Burkina Faso",
    fill = 'toself'
    )%>%
    layout(
        polar = list(radialaxis = list(visible = T,
                                       range = c(0,100)) )
    )

figKE<-plot_ly(
    type = 'scatterpolar',
    r = c(40,43,70,28,73,43,62,40),
    theta = c("FP.agency", "FP.efficacy", 
              "Information", "Affordability.insurance", "Method.choice",
              "Quality1", "Quality2",
              "FP.agency"),
    name="Kenya",
    fill = 'toself'
    )%>%
    layout(
        polar = list(radialaxis = list(visible = T,
                                       range = c(0,100)) )
    )

subplot(figBF, figKE, nrows=1)
```

```{r, eval=FALSE}

#from: https://stackoverflow.com/questions/37670412/can-i-recreate-this-polar-coordinate-spider-chart-in-plotly/37778091#37778091


d <- structure(list(Year = rep(c("2015","2016"),each=24),
                    Response = structure(rep(1L:24L,2), 
                                         .Label = c("Trustworthy", "Supportive", "Leading",
                                                    "Strong", "Dependable", "Consultative",
                                                    "Knowledgeable", "Sensible", 
                                                    "Intelligent", "Consistent", "Stable", 
                                                    "Innovative", "Aggressive", 
                                                    "Conservative", "Visionary", 
                                                    "Arrogant", "Professional", 
                                                    "Responsive", "Confident", "Accessible", 
                                                    "Timely", "Focused", "Niche", "None"),
                                         class = "factor"), 
                    Proportion = c(0.54, 0.48, 0.33, 0.35, 0.47, 0.3, 0.43, 0.29, 0.36,
                                   0.38, 0.45, 0.32, 0.27, 0.22, 0.26,0.95, 0.57, 0.42, 
                                   0.38, 0.5, 0.31, 0.31, 0.12, 0.88, 0.55, 0.55, 0.31,
                                   0.4, 0.5, 0.34, 0.53, 0.3, 0.41, 0.41, 0.46, 0.34, 
                                   0.22, 0.17, 0.28, 0.94, 0.62, 0.46, 0.41, 0.53, 0.34, 
                                   0.36, 0.1, 0.84), 
                    n = rep(c(240L,258L),each=24)
                    ),
               .Names = c("Year", "Response", "Proportion", "n"), 
               row.names = c(NA, -48L), 
               class = c("tbl_df", "tbl", "data.frame"))

str(d)
dim(d)

### 
scale <- 10   # multiply latitudes by a factor of 10 to scale plot to good size in initial view
d$lat <- scale*d$Proportion - 90
d$long <- (as.numeric(d$Response)-1) * 360/24

###
p <- plot_ly(d[c(1:24,1,25:48,25),], lat=lat, lon=long, 
             color = factor(Year), colors=c('#F8756B','#00BDC2'),
             type = 'scattergeo', mode = 'lines+markers', showlegend=T) %>%
    layout(geo = list(scope='world', showland=F, showcoastlines=F, showframe=F,
                      projection = list(type = 'azimuthal equidistant', rotation=list(lat=-90), scale=5)),
           legend=list(x=0.7,y=0.85))
p

###
p %<>% add_trace(type="scattergeo",  mode = "text", lat=rep(scale*1.1-90,24), lon=long, 
                 text=Response, showlegend=F, textfont=list(size=10)) %>%
       add_trace(type="scattergeo",  mode = "text", showlegend=F, textfont=list(size=12),
                 lat=seq(-90, -90+scale,length.out = 5), lon=rep(0,5), 
                 text=c("","25%","50%","75%","100%"))

###
l1 <- list(width = 0.5, color = rgb(.5,.5,.5), dash = "3px")
l2 <- list(width = 0.5, color = rgb(.5,.5,.5))
for (i in c(0.1, 0.25, 0.5, 0.75, 1)) 
    p <- add_trace(lat=rep(-90, 100)-scale*i, lon=seq(0,360, length.out=100), type='scattergeo', mode='lines', line=l1, showlegend=F, evaluate=T)
for (i in 1:24) 
    p <- add_trace(p,lat=c(-90+scale*0.1,-90+scale), lon=rep(i*360/24,2), type='scattergeo', mode='lines', line=l2, showlegend=F, evaluate=T)

```
